---
title: "Digital Inequality in U.S. Broadband Access"
format: html
execute: 
  echo: false
---


```{r}
#| echo: false
#| warning: false
# Load required packages
library(tidyverse)
library(tidycensus)
# Load necessary packages
library(tidyverse)

# Load required libraries
library(plotly)
library(readr)
library(dplyr)

# Load cleaned data
acs_clean <- read_csv("acs_clean.csv")

```

```{r}
# Linear model
model <- lm(broadband_total ~ median_income_k + pct_white + pct_bachelor + pct_enrolled + median_age,
            data = acs_clean)
```



```{r}
#| echo: false
#| warning: false
# Predicted values for each county using the model
acs_clean$predicted <- predict(model, newdata = acs_clean)

```

```{r}
# 📊 Create coef_df for lollipop plot
conf_int <- confint(model)
coef_df <- broom::tidy(model) %>%
  filter(term != "(Intercept)") %>%
  mutate(lower = conf_int[term, 1],
         upper = conf_int[term, 2],
         effect = ifelse(estimate > 0, "Positive", "Negative"),
         variable = case_when(
           term == "median_income_k" ~ "Median Income",
           term == "pct_white" ~ "% White",
           term == "pct_bachelor" ~ "% Bachelor's Degree",
           term == "pct_enrolled" ~ "% Enrolled in School",
           term == "median_age" ~ "Median Age",
           TRUE ~ term
         ))

```

```{r}
ggplot(coef_df, aes(x = reorder(variable, estimate), y = estimate)) +
  geom_segment(aes(xend = variable, y = 0, yend = estimate, color = effect), 
               linewidth = 1.2, lineend = "round") +
  geom_point(aes(color = effect), size = 6) +
  geom_errorbar(aes(ymin = lower, ymax = upper, color = effect), width = 0.15, linewidth = 1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40", linewidth = 1) +
  coord_flip() +
  scale_color_manual(values = c("Positive" = "#2C7BB6", "Negative" = "#D7191C")) +
  labs(
    title = "Digital Inequality in U.S. Broadband Access",
    subtitle = "Effect of Socioeconomic Variables on Broadband Availability(95% CI)",
    x = "Predictor Variable",
    y = "Effect on Broadband Access (%)",
    color = "Effect Direction"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 12, hjust = 0),
    plot.subtitle = element_text(size = 08, hjust = 0, margin = margin(b = 10)),
    legend.position = "top",
    axis.title.y = element_blank()
  )



```

Broadband internet access across U.S. counties varies significantly with income, race, age, and education levels. Using data from the U.S. Census Bureau’s American Community Survey (ACS) 5-year estimates for 2023, I examined county-level variation in broadband access in relation to median household income, racial composition, educational attainment, school enrollment, and median age. The data were retrieved using the tidycensus R package and processed to create a clean, wide-format dataset containing one row per county and five key socioeconomic variables. Broadband access was modeled as a linear function of these predictors. The results show that counties with higher income levels and larger White populations tend to have significantly greater broadband access, even after accounting for other factors. In particular, a $1,000 increase in median household income is associated with a 2.34% increase in broadband access (95% CI: 2.06%, 2.61%). A one percentage point increase in the White population predicts a 0.42% increase in broadband access (95% CI: 0.01%, 0.83%). On the other hand, each 1% increase in adults with a bachelor’s degree is associated with a 2.35% decrease in broadband access (95% CI: –3.38%, –1.32%), while school enrollment is associated with a 3.06% decrease (95% CI: –4.01%, –2.12%), and median age shows a negative effect of –1.10% per additional year (95% CI: –2.00%, –0.22%). These findings suggest that disparities in broadband coverage are influenced not only by economic resources but also by complex demographic and structural factors that may disadvantage certain communities. An interactive visualization further illustrates these relationships, clearly depicting the positive effects of income and race, and the negative effects of education, school enrollment, and age, on predicted broadband access across U.S. counties.